What paradigms can webcam eye-tracking be used for? Attempted replications of five ``classic'' cognitive science experiments

### 2024-08-22 group-e/E5-analysis-calibration.Rmd
E5 edits (wordsmithing, adding back mystery exclusions)

# Excluding Subjects
all.data <- all.data %>% filter(subject_id!="603ea6a51c38def3e4140ba6" & subject_id!="5ce1372c55fe0d001869c8d8" & subject_id!="5f47d860b56e68220cb86be1" & subject_id!="607eaf87ac02229ef4fa8f4a" & subject_id!="5d30857eb2f0ad00163470f3" & subject_id!="5e272684f205790453c2c4b8" & subject_id!="5f5b5881b17fa72756b8ff2a" & subject_id!="5fae5291ec402509e2885604")
E5.excluded.n = 8

@@

subject.calibration.score <- all.data %>%
  filter(trial_type == "webgazer-validate") %>%
  select(subject_id, trial_index, phase,  percent_in_roi, average_offset) %>% 
  unnest(percent_in_roi) %>%
  group_by(subject_id, trial_index, phase) %>% 
  summarize(mean_percent_in_roi = mean(percent_in_roi)) %>% 
  group_by(subject_id) %>% 
  mutate(calib_num = row_number()) %>%
  group_by(subject_id) %>%
  filter(calib_num == max(calib_num)) %>%
  select(subject_id, mean_percent_in_roi)

@@

As in the previous experiments, calibration score was defined as the average proportion of samples within 200 pixels of the validation point during the final validation phase before the eye tracking is performed. The distribution across participants is shown in Fig. \@ref(fig:E5cal). 
(ref:E5cal-caption) Histogram of calibration success in Exp. 5. Where participants required more than one calibration (N=8), only the final calibration was considered.

```{r E5cal, fig.show="hold", out.width="45%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5cal-caption)"}
ggplot(subject.calibration.score, aes(x=mean_percent_in_roi)) +
  stat_bin(binwidth = 10) +
  labs(x="Calibration Accuracy\n(% of samples within 200px of calibration point)") +
  theme_classic()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank())
```

To determine whether calibration accuracy influenced our key effects, we calculated the percentage of samples during the task in which the participant was fixating the face they ultimately chose. There was a significant correlation for both the attractiveness judgments (r = `r round(scorescorrs.A$estimate, 2)` [`r round(scorescorrs.A$conf.int, 2)[1]`, `r round(scorescorrs.A$conf.int, 2)[2]`], p = `r round(scorescorrs.A$p.value, 2)`) and the roundness judgments (r = `r round(scorescorrs.R$estimate, 2)` [`r round(scorescorrs.R$conf.int, 2)[1]`, `r round(scorescorrs.R$conf.int, 2)[2]`], p = `r round(scorescorrs.R$p.value, 2)`). Inspection of Fig. \@ref(fig:E5calcorr) reveals that this correlation is due to a handful of particpiants with calibration values below 50%.

(ref:E5calcorr-caption) Correlation between calibration accuracy (x-axis) and percentage of samples fixating target (y-axis) in Exp. 5.

```{r E5calcorr, fig.show="hold", out.width="80%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5calcorr-caption)"}
ggplot(scores, aes(x=mean_percent_in_roi, y=prop.in.target.subject, group=task, color=task))+
  geom_point() + geom_smooth(method="lm") +
  labs(x="calibration accuracy", y="looking at chosen face") +
  theme_classic() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank())
```

>>>>

# Excluding Subjects
# all.data <- all.data %>% filter(subject_id!="603ea6a51c38def3e4140ba6" & subject_id!="5ce1372c55fe0d001869c8d8" & subject_id!="5f47d860b56e68220cb86be1" & subject_id!="607eaf87ac02229ef4fa8f4a" & subject_id!="5d30857eb2f0ad00163470f3" & subject_id!="5e272684f205790453c2c4b8" & subject_id!="5f5b5881b17fa72756b8ff2a" & subject_id!="5fae5291ec402509e2885604")
# E5.excluded.n = 8

#There's an issue with Subject 603ea6a51c38def3e4140ba6 (roundness condition): their webgazer_data is a list of length 0 instead of a dataframe like the others.
all.data <- all.data %>% filter(subject_id!="603ea6a51c38def3e4140ba6")

@@

subject.calibration.score <- all.data %>%
  filter(trial_type == "webgazer-validate") %>%
  select(subject_id, trial_index, phase,  percent_in_roi, average_offset) %>% 
  unnest(percent_in_roi) %>%
  group_by(subject_id, trial_index, phase) %>% 
  summarize(mean_percent_in_roi = mean(percent_in_roi)) %>% 
  group_by(subject_id) %>% 
  mutate(calib_num = row_number()) %>%
  group_by(subject_id) %>%
  filter(calib_num == max(calib_num)) %>%
  select(subject_id, mean_percent_in_roi,calib_num)

@@

As in the previous experiments, calibration score was defined as the average proportion of samples within 200 pixels of the validation point during the final validation phase before the eye tracking is performed. <!--The distribution across participants is shown in Fig. \@ref(fig:E5cal). 
(ref:E5cal-caption) Histogram of calibration success in Exp. 5. (AJ: I cut this to be more similar to the other expts)-->Where participants required more than one calibration (N=14), only the final calibration was considered.

```{r E5cal, fig.show="hold", out.width="45%", echo=FALSE,eval=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5cal-caption)"}
ggplot(subject.calibration.score, aes(x=mean_percent_in_roi)) +
  stat_bin(binwidth = 10) +
  labs(x="Calibration Accuracy\n(% of samples within 200px of calibration point)") +
  theme_classic()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank())
```

To determine whether calibration accuracy influenced our key effects, we calculated the proportion of samples during the task in which the participant was fixating the face they ultimately chose. Calibration accuracy significantly correlated with fixations in both the attractiveness condition (r = `r round(scorescorrs.A$estimate, 2)` [`r round(scorescorrs.A$conf.int, 2)[1]`, `r round(scorescorrs.A$conf.int, 2)[2]`], p = `r round(scorescorrs.A$p.value, 2)`) and the roundness condition (r = `r round(scorescorrs.R$estimate, 2)` [`r round(scorescorrs.R$conf.int, 2)[1]`, `r round(scorescorrs.R$conf.int, 2)[2]`], p = `r round(scorescorrs.R$p.value, 2)`). Inspection of Fig. \@ref(fig:E5calcorr) reveals that this correlation is due to a handful of participants with calibration values below 50%.

(ref:E5calcorr-caption) Correlation between calibration accuracy (x-axis) and proportion of samples fixating target (y-axis) in Exp. 5.

```{r E5calcorr, fig.show="hold", out.width="80%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5calcorr-caption)"}
ggplot(scores, aes(x=mean_percent_in_roi, y=prop.in.target.subject, group=task, color=task))+
  geom_point() + geom_smooth(method="lm") +
  labs(x="Calibration Accuracy", y="Proportion Fixating Chosen Face") +
  #theme_classic() +
  theme_bw() +
  theme(#panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        #axis.title.y = element_blank()
        )
```

### 2024-08-22 manuscript.Rmd
E5 edits (wordsmithing, adding back mystery exclusions)

50 participants for the main task were recruited on Prolific and were
paid \$10/hour. 8 subjects, 4 from the attractiveness task group and 4 from
the roundness task group, were excluded for incorrect validations.
<!-- they were able to proceed despite bad validation?--> After this
data exclusion, we ended up with 21 participants each for the
attractiveness task and the roundness task. The original sample size in
Shimojo et al. (2003) was 10 participants total.

@@

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure\ \@ref(fig:E5-example-trial)). Participants were randomly assigned to one of two tasks:
attractiveness or shape judgment. In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the "a" key on their keyboard to
select the face on the left and the "d" key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials. The order of
the 19 face pairs was random for each participant.

```{r E5-example-trial, fig.cap='An example of a critical trial from Experiment 5. (Text did not appear on each screen.)'}
knitr::include_graphics("group-e/E5-example-figure.jpeg")
```

### Materials and Norming

The faces in our replication were selected from a set of 1,000 faces
within the Flickr-Faces-HQ Dataset. (The face images used in Shimojo et
al. were from the Ekman face database and the AR face database.) These
images were chosen because the person in each image was looking at the
camera with a fairly neutral facial expression and appeared to be over
the age of 18. 27 participants were recruited on Prolific to participate
in stimulus norming (for attractiveness). They each viewed all 172 faces and were asked to rate
them on a scale from 1 (less attractive) to 7 (more attractive) using a
slider. Faces were presented one at a time and in a random order for
each participant. Data from 3 participants were excluded because
their mode response made up more than 50% of their total responses, for a total of 24 participants
in the norming.

In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position was then determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz.`r colorize("[TODO - CONFIRM]", 'red')`

>>>>

Fifty participants for the main task were recruited on Prolific and were
paid \$10/hour. <!-- Eight subjects, 4 from the attractiveness task group and 4 from
the roundness task group, were excluded for incorrect validations.
(They were able to proceed despite bad validation?) After this
data exclusion, we ended up with 21 participants each for the
attractiveness task and the roundness task. AJ: I can't figure out why these 8 were excluded so I'm adding them back for now--> For one subject (roundess task group), eye gaze data failed to record.<!--This was one of the excluded 8--> We ended up with 25 participants in the attractiveness condition and 24 in the roundness condition. The original sample size in Shimojo et al. (2003) was 10 participants total.

The faces in our replication were selected from a set of 1,000 faces
within the Flickr-Faces-HQ Dataset (the face images used in Shimojo et
al. were from the Ekman face database and the AR face database). These
images were chosen because the person in each image was looking at the
camera with a fairly neutral facial expression and appeared to be over
the age of 18. Twenty-seven participants were recruited on Prolific to participate in stimulus norming (for attractiveness). They each viewed all 172 faces and were asked to rate them on a scale from 1 (less attractive) to 7 (more attractive) using a slider. Faces were presented one at a time and in a random  order for each participant. Data from three participants were excluded because
their modal response made up more than 50% of their total responses, for a total of 24 participants in the norming.

Participants were presented with each of the 19 face pairs; task condition was manipulated between subjects, such that half of participants made judgments about facial attractiveness and the other half made judgments about facial shape.

@@

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure\ \@ref(fig:E5-example-trial)). In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the "a" key on their keyboard to
select the face on the left and the "d" key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials. The order of
the 19 face pairs was random for each participant.

```{r E5-example-trial, fig.cap='An example of a critical trial from Experiment 5 (text did not appear on each screen).'}
knitr::include_graphics("group-e/E5-example-figure.jpeg")
```

## Results
In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz. <!--`r colorize("[TODO - CONFIRM]", 'red')` AJ: I don't know how to confirm this! -->
